---
title:           Initial exploration of Sleep apneoa 
author:          Dennis Wollersheim 
date:            11.07.2018
linkcolor:       cyan
citecolor:       grey
output:
  pdf_document:
    highlight:   zenburn
---

\tableofcontents

# motivation 
We want to investigate the sleep apnea pp, and to do this, we need to have something to compare them to.  We will create a list of patients, consisting of all the sleep apnea people, and then also, people who are similiar to the sleep apnea people in age, gender, and sex, who have not had a sleep apnea test. 


```{r results='hide', message=FALSE, warning=FALSE}

source('lib/functions.R')
source( "lib/get_data.R")
# -------------------------------------------------

my_db_get_query( 'select * from mbs_apnea_unique' ) %>%
  mutate( type='apnea') %>%
as.tibble() %>%
{.} -> ap


my_db_get_query( 'select * from mbs_apnea_matched_sample ' ) %>%
  anti_join( ap, by='pin') %>%
  mutate( type='non-apnea') %>%
  as.tibble() %>%
  {.} -> apm

ap %>%
  select( pin, sex, yob, nservices, type ) %>%
  bind_rows( apm ) %>%
  mutate( scat = cut_number( nservices, 5  ),
         yob_group = cut_number( yob, 5 )) %>%
  {.} -> apboth

```

how many of each type of person are there

```{r}
apboth %>%
  count( type )

```


# initial exploration
## what does the apnea dataset look like
 * Take all pp who had apnea, break them into gender, birth year

 * Take everyone else who had the same gender and birthyear  

 * count the number of services for each person took over the 20 year period

 * note: the apnea people are generally larger service users than the non-apnea pp


```{r}

apboth %>%
  ggplot() +
  geom_histogram( aes( nservices, fill=type)) +
  scale_x_log10() + 
  facet_grid( type ~ sex, scales="free_y" )

```
## match apnea with non-apnea
Due to this disparity, we match each apnea person with a simliar person from from the non-apnea set, who had the same number of services, and use them.

First, make sure that every apnea has a match in non-apnea. 1) get the target, and 2) make sure that the non-apnea has each of the target items

```{r}

apboth %>%
  filter( type=="apnea") %>%
  count( sex, yob_group, scat ) %>%
  rename( sample_size = n ) %>% 
  { . } -> target_sample 

apboth %>%
  filter(type=='non-apnea') %>%
  distinct( sex, yob_group, scat ) %>%
  anti_join( target_sample)

```

Now, make sure that there are sufficient non-apnea to match all apnea

```{r}

apboth %>%
  filter(type=='non-apnea') %>%
  count( sex, yob_group, scat ) %>%
  rename( n_non = n ) %>%
  inner_join( target_sample) %>%
  mutate( diff = n_non - sample_size  ) %>%
  filter( diff < 0 )

```

Now, extract matching non-apnea

```{r}

apboth %>%
  filter(type=='non-apnea') %>%
  nest( -sex, -yob_group, -scat ) %>%
  inner_join( target_sample) %>%
  mutate( sample = map2( data, sample_size , sample_n )) %>%
  unnest(sample) %>%
  bind_rows( ap ) %>% 
  { . } -> ap_balanced

ap_balanced

 sample


```

### Note that number of services is now balanced between the groups

```{r}

ap_balanced %>%
ggplot() +
  geom_histogram( aes( nservices, fill=type)) +
  scale_x_log10() + 
  facet_grid( type ~ sex)

ap_balanced %>% write.csv("output/ap_balanced.csv")

```



